---
title: "Homework 24"
output: html_notebook
---

# Introduction

This assignment will walk through how to use R for statistical tests.
With each statistical test, the following workflow will be followed:
1. Plot your data
2. Build your model
3. Check your assumptions
4. Interpret your model
5. Replot the data and the model

## Getting Set Up

```{r}
rm(list = ls())
library(dplyr)
library(ggplot2)
library(here)
```

## Pulling in the data
The first dataset being used is called 'ladybirds'.
We will use it to run a Chi-squared contingency analysis.

```{r}
lady <- read.csv(here("Data", "ladybirds.csv"), stringsAsFactors = TRUE) 
glimpse(lady) #checks out dataset
```

## Grouping the data for use
The groupby() and summarise() functions are best for sorting the data into groups that we would like to explore further.

```{r}
totals <- lady %>%
  group_by(Habitat, colour) %>%
  summarise(total.number = sum(number))
```

## Generating the first figure 


```{r}
ggplot(totals, aes(x = Habitat, y = total.number, 
                   fill = colour)) + 
  geom_bar(stat = 'identity', position = 'dodge') #identity tells ggplot to not try to calculate anything from the data.
#dodge tells R to provide two bars next to each other under each variable
```

### Fixing the colors

There is a way to customize the colors of the bars. In this case we can make a bar black to match the black individuals in the dataset and the other bar can be red and correspond to the red individuals.

```{r}
ggplot(totals, aes(x = Habitat, y = total.number, 
                   fill = colour)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_manual(values = c(black = "black", red = "red"))
```

## Running a Chi-squared test

```{r}
lady_mat <- xtabs(number ~ Habitat + colour,
                  data = lady)
lady_chi <- chisq.test(lady_mat)
names(lady_chi)
lady_chi$expected
```

## Running a T-test

### Clear R's brain, activate packages and pull in correct dataset

```{r}
ozone <- read.csv(here("Data", "ozone.csv"))
glimpse(ozone)
```

## Plot the data

```{r}
ggplot(ozone, aes(x = Ozone)) +
  geom_histogram(binwidth = 10) +
  facet_wrap(~ Garden.location, ncol = 1) +
  theme_bw()
```

## Grouping the data

```{r}
totals <- ozone %>%
  group_by(Garden.location, Garden.ID) %>%
  summarise(total.number = sum(Ozone))
```

## Run a T-Test

```{r}
t.test(Ozone ~ Garden.location, data = ozone)
```

Another potential test to run:
```{r}
var.test(Ozone ~ Garden.location, data = ozone)
```

