---
title: "GSWR Ch4"
output: html_notebook
---

## Introduction
This assignment walks through Chapter 4: Visualizing Your Data in the Getting Started With R textbook.

### Getting Started
Generate code to activate the correct packages and clear R's brain.

```{r}
rm(list = ls())
library(dplyr)
library(ggplot2)
library(here)
```

Pull in data

```{r}
Compensation <- read.csv(here("Data", "compensation.csv"), stringsAsFactors = TRUE)
```

### Practice with scatterplots

```{r}
ggplot(Compensation, aes(x = Root, y = Fruit)) + geom_point() #creates a basic scatterplot
```

### Customizing plots
With the basic plot generated above, the background is grey. One way to remove this is with there theme_bw() function.

```{r}
ggplot(Compensation, aes(x = Root, y = Fruit)) + geom_point() + theme_bw() #generates a scatterplot without a grey background.
```

The size of the points can be changed by adding a value to the function geom_point() 

```{r}
ggplot(Compensation, aes(x = Root, y = Fruit)) + geom_point(size = 5) + theme_bw()
```

The axis labels can be altered using xlab() and ylab()

```{r}
ggplot(Compensation, aes(x = Root, y = Fruit)) + geom_point(size = 5) + xlab("Root Biomass") + ylab("Fruit Production") + theme_bw()
```

The colors of points that match a specific group can be adjusted to represent levels with the group.

```{r}
ggplot(Compensation, aes(x = Root, y = Fruit, colour = Grazing)) + geom_point(size = 5) + xlab("Root Biomass") + ylab("Fruit Production") + theme_bw()
```

Similarly to the color changing, the shape of the points can also be changed using the same format as the previous graph.

```{r}
ggplot(Compensation, aes(x = Root, y = Fruit, shape = Grazing)) + geom_point(size = 5) + xlab("Root Biomass") + ylab("Fruit Production") + theme_bw()
```

### Practice with Box-and-whisker plots 

Setting up a basic box plot

```{r}
ggplot(Compensation, aes(x = Grazing, y = Fruit)) + geom_boxplot() + xlab("Grazing treatment") + ylab("Fruit Production") + theme_bw()
```

Box plots can be customized 

```{r}
ggplot(Compensation, aes(x = Grazing, y = Fruit)) + geom_boxplot() + geom_point(size = 4, colour = 'lightgrey', alphs = 0.5) + xlab("Grazing treatment") + ylab("Fruit Production") + theme_bw()
```

### Practice with Histograms

Generate histograms of fruit production using the generic composition

```{r}
ggplot(Compensation, aes(x = Fruit)) + geom_histogram()
```

The result of the generic histogram does not fit the dataset well and is visually unappealing. It can be fixed by specifying the bin width.

```{r}
ggplot(Compensation, aes(x = Fruit)) + geom_histogram(binwidth = 15)
```

You can also generate more than one histogram using the facet_wrap() function

```{r}
ggplot(Compensation, aes(x = Fruit)) + geom_histogram(binwidth = 15) + facet_wrap(~Grazing)
```

Woohoo Graphs!