---
title: "Sirois_Homework12.Rmd"
author: "Julia Sirois"
date: "9/27/2022"
output: html_document
---

During this homework, students are required to work with a new dataset and apply new skills learned in RStudio.

The first step is to set up.

```{r}
rm(list = ls()) #clears R's brain
library(dplyr) #library() turns on the specified packages
library(ggplot2)
library(here)
```

Get the data 

```{r}
Cover <- read.csv(here("Data", "Cover.csv"), stringsAsFactors = TRUE) #pulls in and names the data
summary(Cover) #gives us a quick summary of the data.
```

Check out the data using different functions

```{r}
names(Cover) #pulls up the column headings
slice(Cover, c(2,9,12,20)) #pulls up specified rows of data
```

The average percent of cover was determined to be 77.5% for the Bare type.

```{r}
Bare <- filter(Cover, Site == "Donnerville 1", Type == "Bare")
summary(Bare)
```


Next, the average percent cover was determined to be 2.5% for the Bryophytes type using the same coding process.

```{r}
Bryophytes <- filter(Cover, Site == 'Donnerville 1', Type == 'Bryophytes')
summary(Bryophytes)
```

Finally, the average percent determined to be 20% for the Vascular type using the same process as mentioned before.

```{r}
Vascular <- filter(Cover, Site == 'Donnerville 1', Type == 'Vascular')
summary(Vascular)
```

The Vascular plant cover percentage was also calculated for all sites. The vascular cover is 16.67% on average across all sites.

```{r}
Vascular_all <- filter(Cover, Type == "Vascular")
summary(Vascular_all)
```

Work to get comfortable with `group_by()` and `summarize()`.  Those are what you need to more efficiently answer the requested questions. Your way worked, but did not result in values that were saved - rather, the user has to read the output of the `summary()` call. 

Here's an example:Determine the average percent cover by cover type (bare, bryophytes, vascular plants) for Donnerville 1.

```{r}
Cover %>% filter(Site == "Donnerville 1") %>% group_by(Type) %>% summarise(avg_percent_cover = mean(Percent, na.rm = T))
```
See if you can use similar to do step 10 from this homework.
