---
title: "Fixing mistakes with dplyr"
author: "Julia Sirois"
date: "10/2/2022"
output: html_document
---

## Introduction
This assignment is intended to help practice the functions in the dpylr package and how to fix errors when they arise

### Getting started 
Turning on the correct packages and clearing R's mind is the first step.

```{r}
rm(list = ls()) #clears r's brain
library(dplyr) #turns on package
library(here)
```

### Pulling in data
Use the here() function to pull in the correct data file. Do not use stringsAsFactors. 

```{r}
DF <- read.csv(here("Data", "drug_resistance.csv")) #Assigns the title 'DF' to the dataframe and pulls in the correct dataset.
summary(DF) #provides a summary of 'DF'
```

Now upload the data again but with stringsAsFactors = TRUE.

```{r}
DF <- read.csv(here("Data", "drug_resistance.csv"), stringsAsFactors = TRUE)
summary(DF)
```
Usings stringsAsFactors = TRUE causes the summary() function to read the qualitative inputs as valid variables whereas without stringsAsFactors = TRUE the summary() function just summarizes them as characters.

### Using other functions
Using the levels() function, pulls the levels attribute for a variable.

```{r}
levels(DF$Treatment)
levels(DF$Plate_n)
levels(colony.count)
levels(tech.name)
```
There are errors for 'colony.count' and 'tech.name'. This is likely due to the '.' in the column headings which indicates there is a space within it.
Change the column names may fix the issue.

```{r}
colnames(DF) <- c("Treatment", "Plate_num", "Colony_count", "Tech_name") #changes the column headings
names(DF) #pulls up the column headings
```

Now that the column headings has been fixed, run the levels() function again to see if the error still occurs.

```{r}
levels(DF$Treatment)
levels(DF$Plate_num)
levels(DF$Colony_count)
levels(DF$Tech_name)
```

The error has been resolved. 