---
title: "HW 25"
output: html_notebook
---

# Introduction
This assignment works through how linear regressions are use in R.

## Getting set up
Activate necessary packages and the function to clear R's brain.

```{r}
rm(list = ls())
library(here)
library(ggplot2)
library(dplyr)
```

## Pull in the data

```{r}
plant_gr <- read.csv(here("Data", "plant.growth.rate.csv"))
```

## Check out the data

```{r}
glimpse(plant_gr)
```

## Generate a scatterplot

```{r}
ggplot(plant_gr,
       aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  xlab("Soil moisture content") +
  ylab("Plant Growth Rate (mm/week") +
  theme_bw()
```
WHAT DO YOU SEE FROM THE SCATTERPLOT? CAN YOU GUESS THE SLOPE AND Y-INTERCEPT BEFORE YOU RUN THE MODEL? WHY IS A LINEAR REGRESSION THE APPROPRIATE TEST.

## Create the first linear model

```{r}
model_pgr <- lm(plant.growth.rate ~ soil.moisture.content,
                data = plant_gr) #hypothesizes that plant growth is a function of soil moisture.
```

## Activate new package and plot the lm

```{r}
library(ggfortify)
autoplot(model_pgr, smooth.colour = NA)
```
WHAT DOES THIS TELL YOU? YOU SHOULD MAKE SOME NOTES
## Interpret the data with an ANOVA

```{r}
anova(model_pgr) #runs the ANOVA
```

```{r}
summary(model_pgr) #large F-value indicates that the error variance is small
```

## Plotting a linear regression

```{r}
ggplot(plant_gr, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ylab("Plant Growth Rate (mm/week)") +
  theme_bw()
```

## Analysis of Variance: One-way ANOVA
A different data set is being used for this piece. Clear R's brain and pull in a new data set.

## Pulling in the data

```{r}
daphnia <- read.csv(here("Data", "Daphniagrowth.csv"))
glimpse(daphnia)
```

## Visualize the data

```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) +
  geom_boxplot() +
  theme_bw() +
  coord_flip() #flipping the coords lists out the specific parasites
```
AGAIN, WHAT DO YOU SEE FROM THE PLOT AND WHAT MIGHT YOU GUESS THE OUTCOME OF YOUR STATISTICAL TEST TO BE BEFORE YOU RUN IT?

## Setting up the one-way ANOVA

```{r}
model_grow <- lm(growth.rate~parasite, data = daphnia)
library(ggfortify) #activates package
autoplot(model_grow, smooth.colour = NA) #checks assumptions
```

## Create the one-way ANOVA

```{r}
anova(model_grow) #has a significant result
summary(model_grow) #shows intercept and the treatments
```

## Get the means of each treatment

```{r}
sumDat <- daphnia %>%
  group_by(parasite) %>%
  summarise(meanGR = mean(growth.rate))
```

YOUR CODE IS ALL FINE BUT I CAN'T TELL IF YOU ARE THINKING ABOUT WHAT YOU ARE DOING BECAUSE YOU HAVEN'T PROVIDED MUCH ANNOTATION TO THE CODE.