---
title: "Practicing dplyr"
author: "Julia Sirois"
date: "10/12/2022"
output: html_document
---

# Introduction
The following exercise is to revisit functions learned in Chapters 2 and 3.

Getting set up
```{r}
rm(list= ls()) #clears R's Brain
library(dplyr) #library() actives the specified package
library(here)
library(ggplot2)
```

Pulling in the data

```{r}
Leaves <- read.csv(here("Data", "Class_leaf_measurements.csv"), stringsAsFactors = TRUE) 
```

```{r}
str(Leaves) #pulls the structure of the data
```

## Cleaning up the data

```{r}
names(Leaves) #pulls the column headings
```
Fix errors (spaces and special characters) in column headings

```{r}
colnames(Leaves) <- c("Tree_num", "Species", "Leaf_num", "DBH_cm", "Petiole_length", "Total_length_cm", "Middle_lobe_length_cm", "Middle_lobe_width_cm", "Person") #reassigns column headings
```

Fix inconsistencies the species column

There should only be three species options, however, there is inconsistent data formatting within the species columns. 

Valid inputs will be Red Maple, Sugar Maple, and Norway Maple

```{r}
str(Leaves) #Look at the structure again
Leaves$Species = as.factor(Leaves$Species) #converts to categorical values
levels(Leaves$Species) #pulls the levels for the specified column

```
Now fix the miss spellings and improper capitalizations to give the column 'Species' only 3 levels

```{r}
levels(Leaves$Species) <- list("Norway Maple" = "Norway maple", "Norway Maple" = "Norway mapel", "Norway Maple" = "Norway Maple", "Red Maple" = "red maple", "Red Maple" = "Red maple", "Red Maple" = "Red Maple", "Sugar Maple" = "sugar maple", "Sugar Maple" = "Sugar maple", "Sugar Maple" = "Sugar Maple") #groups the data into the desired levels
levels(Leaves$Species) #check to see if it worked
```

I LIKE THE WAY YOU FIXED THE ERRORS BUT I ALSO WANT TO MAKE SURE YOU FEEL COMFORTABLE USING THE `which()` FUNCTION. TAKE A LOOK AT THE KEY FOR EXAMPLE. 

With the data fixed, subsets can be created

```{r}
Sugar <- filter(Leaves, Species == "Sugar Maple") #creates a subset 
Sugar <- select(Sugar, -Person, -Leaf_num, -Tree_num)
```

Now create a histogram of the petiole lengths from the Sugar subset

```{r}
ggplot(Sugar, aes(x = Petiole_length)) + geom_histogram()
```
LOOKS GOOD. BE CAREFUL ABOUT PUTTING THE + geom_histogram() ON THE SAME LINE - SOMETIMES YOU RUN INTO WEIRD ERRORS IF THE PLUS ISN'T ON THE LINE ABOVE THE NEXT GGPLOT LAYER.
