---
title: "Sirois_HW11_GSWR3Ex1.Rmd"
author: "Julia Sirois"
date: "9/26/2022"
output: html_document
---

### Introduction

In this exercise we will be practicing new functions using a compensation dataset focuses on the production of fruit. 

First, let's set things up

```{r}
rm(list = ls())
library(dplR)
library(ggplot2)
library(here)
```

Now let's pull in our data!

```{r}
DF <- read.csv(here("Data", "compensationdataset.csv"), stringsAsFactors = TRUE)
glimpse(DF) #checking the dataset
```

Now let's get a summary for the compensastion variables.

```{r}
summary(DF)
```

We are now going to practice subsetting.

```{r}
names(DF)
select(DF, Fruit) #we are looking at the Fruit column.
```

We can also use the select() function to pull up all columns except one. Let's practice by leaving out the 'Root' column.

```{r}
select(DF, -Root) #that is a minus sign
```

The slice() function grabs rows of data.

```{r}
slice(DF, 2)
```

We can also slice a range of rows by using a :

```{r}
slice(DF, 2:10)
```

You can also collect the row numbers using the function c()

```{r}
slice(DF, c(2, 3, 10))
```

The filter() function is another useful subsetting. We will first use the with() function to ask R to look at the dataframe, do what comes next, and then stop looking. 

```{r}
with(DF, Fruit > 80) #the > is a logical that produces a sequence of TRUE and FALSE.
```

Now let's use filter(). Let's find the rows where it is true that Fruit is >80 return.

```{r}
filter(DF, Fruit > 80)
```

We can easily select rows according to multiple conditions. 

```{r}
filter(DF, Fruit > 80 | Fruit < 20) # | is a symbol that indicates 'or'
```

We need to make sure we can use the subset of our data.

```{r}
lo_hi_fruit <- filter(DF, Fruit > 80 | Fruit < 20) #creates saved subset
lo_hi_fruit #views subset
```

# What does the composition look like now?
```{r}
head(DF) #opens first 6 rows of the new DF
```

We are now going to use mutate()

```{r}
DF <- mutate(DF, logFruit = log(Fruit))
```

Let's look at the first six rows again.

```{r}
head(DF)
```

It is also possible to sort our data. Let's practice by using the arrange() function.

```{r}
arrange(DF, Fruit)
```

Root values from Fruit > 80 subset

```{r}
select(filter(DF, Fruit > 80), Root)
```

There are multiple ways to summerize our data. Let's go over a couple of them.

Method 1: Nested, no pipe
```{r}
summarise(group_by(DF, Grazing) , meanFruit = mean(Fruit))
```

Method 2: Pipe, no nesting

```{r}
DF %>%
group_by(Grazing) %>%
  summarize(meanFruit = mean(Fruit))
```

Summarizing and extending summarization

```{r}
DF %>%
group_by(Grazing) %>%
  summarise(meanFruit = mean(Fruit), sdFruit = sd(Fruit))
```

