---
title: "Sirois_Ch2_Pt2"
author: "Julia Sirois"
date: "9/18/2022"
output: html_document
---

## Introduction
This analysis looks at a dataset called "nasty format"

### Getting started
The first step is to import the data

```{r}
library(readr) #turns on readr package
nasty_format <- read_csv("Data/nasty format.csv") #specifies the file pathway for the dataset
```

Now we will assign the variable "nasty.format" to the data file.
```{r}
read.csv("Data/nasty format.csv") #reads the file we wish to associate to a variable
```

```{r}
nasty.format <- read.csv("Data/nasty format.csv") #assigns the variable
```

We can now type in the variable we assigned to the data while running different functions.
```{r}
str(nasty.format) #returns several pieces of information about the dataset
```

We can remove unwanted rows of data with the 'filter()' function. 
```{r}
library(dplyr) #turns on dplyr package
nasty.format <- filter(nasty.format, Bottle != "") #filters the data
glimpse(nasty.format) #provides a horizontal view of the data
```

Now we are going to tidy the data. We can do this with the function 'gather()'

```{r}
library(tidyr) #turns on tidyr package
tidy_data <- gather(nasty.format, Date, Abundance, 4:11) #tidies the data based on variables specified in the parentheses.
```

We will now use 'str()' and 'glimpse()' to see if it worked.

```{r}
str(tidy_data) #provides us with information about the dataset
glimpse(tidy_data) #provides horizontal view of data
```

### Cleaning the dates
We will now clean up the dates within the dataset by removing the X by using the function 'sub_str()'



```{r}
library(stringr) #turns on stringr package
tidy_data <- mutate(tidy_data, Date = substr(Date, 2, 20)) #removes unwanted variables from date
```

```{r}
unique(tidy_data$Date) #pulls up date values
```

```{r}
library(lubridate) #turns on lubridate package
tidy_data <- mutate(tidy_data, Date = dmy(Date)) #generates a new date organization method.
```

```{r}
glimpse(tidy_data) #provides horizontal view of data 
```

Now we can generate plots for our tidied data.

```{r}
library (ggplot2) #turns on ggplot2 package
ggplot(data = tidy_data, aes(x=Date, y=Abundance)) + geom_point() + facet_wrap(~Bottle)
```

